A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-03-28 paper-hierarchical.tex
more work

The next step would be to estimate the number of photons at the sample. For that, we need the spectrum at the source. It is not straightforward to get this information ``by hand'' but one can calculate the flux and power using different codes. Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeUS} and SRW \cite{codeSRW}. Figure~\ref{fig:FluxU18} shows the spectra for two slit apertures calculated with SRW. The spectrum is calculated over an aperture opened to fully accept the central cone at the resonance peak of the harmonic in use. placed at $d$=27.1m. An aperture size of $S_h=S_v=4$ mm it is much larger than the Full Width at Half Maximum (FWHM) of the central cone at $d$, that is $\max(\Sigma') d$ \todo{HOW MUCH?}, with any possible value of $\Sigma'$ in Table~\ref{tab:HandCalculations}. The peak intensity is about  $3.5\cdot10^{14}$ photons/s/0.1{\%}bw (see \ref{fig:FluxU18}). The corresponding spectral peak is very wide and depends mostly on the aperture so it does not change when reducing emittance (from high-$\beta$ (ESRF) to EBS). For smaller apertures (1~mm) the emittance plays an important role in both peak width and peak amplitude. The number of photons at the source at photon energy of 17225 eV \inred{(3.5 $10^{14}$ /17.225 photons/s/eV)} has to be multiplied by the bandpass of the monochromator at that energy \inred{(17225 $10^{-2}$)}, thus resulting $N_{ML}$=3.5 $10^{15}$ photons/s.

\begin{figure}\label{fig:FluxU18}
    \centering
    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.jpg}
    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
\end{figure}

\inred{Next we would like to estimate} the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons. This can be analysed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50 \mu$m in horizontal and fully opened in vertical. For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis. Using mirror length of 60 mm and 26 mm for VFM and HFM, respectively, and incident angle of 15 mrad mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$. Dividing this by the distance from the mirror to the VSS in H and to the source in V we get N.A. of6.2$\times$21.6 $mu$rad. Dividing by the beam divergence we have the new transmittivity coefficients 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS. The overall transmission is therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 7.53 $10^{12}$ photons for ESRF and 1.03 $10^{14}$ photons for EBS. \todo{check with referente 6 $10^9$ at 33.6 keV}

One can also discuss some simple consequences about the beam coherence. The coherence fraction $CF$ \cite{arxivCF} of the undulator beam is the occupation of the first coherence mode, as it will be discussed in Section~\ref{level2}. An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}, for the horizontal and vertical planes we have have
\begin{equation}
 CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
\end{equation}
and the total coherent fraction is the product of both ($CF=CF_h \times CF_v$). An estimation of the CF calculated in such a mode gived CF=0.XX\% for the ESRF and XXX for EBS. The CF roughly indicates the ration of the ``coherent photons'' over the total number of photons. A the beamline optics of a beamline target for coherence application must therefore clean the beam to extract only the good coherent photons. A complete cleaning is impossible as there is spatial overlapping of the different coherence modes. However, apertures centered at the optical axis would make the work. If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) we find them very close to the $CF$ values, meaning that the resulting beam would be highly coherent.

Regarding coherence, one can idealize the beamline in two parts: the ML onochromator and VSS that will prepare a coherent plane wave, as the source is very far from the sample position, and then KB that acts as a focusing element (focal=cm) that also crops the beam like it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}. This aperture has two effects: i) removing a lot of off-axis photons what imply a increase of the coherent fraction to a situation close to a coherent source, and ii) make a broadening of the focal spot because of coherent diffraction. We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution of:
\begin{equation}
 I(x,y,z=f) = \frac{4 k^2 D^2}{\pi^2 f^2} \sinc^2\frac{k D x}{2 f} \sinc^2\frac{k D y}{2 f}
\end{equation}
where $k=2\pi/\lambda$ . Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ nm for the focal spot. This spot size is of the order or smaller than the one obtained by geometrical considerations, therefore one can say that this beamline for the conditions in use is a diffraction limited beamline.

In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (XX$\times$XX nm$^2$) the flux (XX photons/s) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$. We ignored at this point the effect of real optics including surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. A similar work was done at the design level in order to define the working parameters.

@@

In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) to the beam size and beamline transmittivity. The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations. The second subsection shows how the effect of slopes erroin the KB mirrors deteriorate the focal spot.

Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the beam characteristics at the image (manily size and divergences). For simple systems one can perform ray tracing by hand, selecting some principal rays that will define the location and envelope of the image. Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics. Rays are usually generated by Monte Carlo sampling the source characteristics. Ray tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along straight line. Rays are intercepted by the optical elements, which change their trajectory. For reflectors, this change of direction is given by the specular reflection laws. For refractors, the change of direction os defined by the Snell law. Because Snell law uses the refraction index which is wavelength-dependent, it is then interetisting to add additional parameters to each ray, such as the wavelength. This will permit to calculate the refraction index that has to be used by each ray. The information of each ray can also be extended to include some intensity, or better electric field amplitude. This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity. For mirrors and lenses one should use the Fresnel's equations for reflection and transmission to affect the ray intensity. This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, CRLs and transfocators; gratings of any type and shape (plane, VLS), crystal systems, etc. It makes ray tracing a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline. Although very powerful commercial ray tracing packages are available in the market, they are not suitable for synchrotron simulations because they do not cover the particularities of X-ray devices and sources. For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}. SHADOW is probably the one that is more used and cited. We use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.

@@

We performed ray tracing using the elliptical reflectors with finite dimensions as focusing elements. The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing. For a first calculation  the elliptical elliptical reflectors are considered perfect, and have no slope errors. 
The results in Fig.~\ref{fig:ray tracing}a show two facts. One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. Secontly, one can observe a much larger luminosity in the EBS case. The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS. These values confirm what we calculated by hand (0.2\% for ESRF and and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.13 and 2.8\%, respectively), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.

@@

The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position. It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Squate (RMS) and then building synthetic surface profiles to be used in the simulations. This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet. However, in our case the mirrors already exist and have been extensively characterize at the ESRF metrology laboratory. It is therefore highly recommended to use metrology data from real mirrors, when possible. When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors. The metrology data of the mirrors for the beamline under study are presented in \todo{Appendix 1}. These experimental data are used in all the simulations in this paper. Simulations using the hybrid model including slopes errors are shown in Fig.~\ref{fig:ray tracing}c. It can be observed a small broadening of the peak due to slope errors, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors. \todo{may be play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?}

@@

\subsubsection{Simplified wave optics simulation (coherent case in 1D with point source and ideal elements)}

The characteristics of this particular beamline, a the very long length combined with a small source (in particular for EBS), makes this beamline very adapted for studied requiring coherence. Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.  As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be sone using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed at 5cm upstream the focus (at the second KB mirror position). The slit vertical size corresponds roughly to the horizontal acceptance of the seconf KB mirror, and it is approximately the same in vertical. A simple simulation of diffraction by a square slit using OASYS is shown in Fig.~\ref{wofry400um}. It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction limited beamline.

\begin{figure}
\label{wofry400um}
\centering
\includegraphics[width=4.25cm]{GRAPHICS/wofry400umslit.png}
\includegraphics[width=4.25cm]{GRAPHICS/wofry400umprofile.png}
\caption{Diffraction spot producedof a spherical wavefront (R=-50mm) by a 400x}
\end{figure}

The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline
The wavefront simulation of the beamline include propagation of the beam from element to element (drift spaces) using different possible propagators. Usually a propagator based on the Freskel-Kirchhoff integral is used, as it is valid for both near and far field. The applications of this (and other) operators imply the calculation of an integral (or sum) at each pixel of the image plane. This is costly from the computer point of view, so Fourier Transforms (Fourier Optics) are used to reduce complexity and calculation time. This is practically the only choice for simulating 2D images. The use of Fourier Transforms adds another problem: the result is very dependent on the sampling of the wavefront (i.e., pixel size, number of points, dimension of the window where the wavefront is sample). This means the usually one has to test many configurations before finding one that works reaonably well (even though one can never guarantee the complete accuracy of the sampling used). A full wavefront optics simulation is an iterative process, where the user has to refine the sampling parameters element by element to be sure the wavefront at each position is sampled correctly. This is a time consuming work for performing it with the full 2D simulation. For this reason, a wavefront calculation may be started by a simplified system, using simplified models for the source and focusing elements. Moreover, because the horizontal and vertical coupling is usually small in synchrotron beamlines, it is convinient to separate the simulation in two, one for the horizontal plane and one for the vertical one.

We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes. The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u D$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $D$ the distance from the point source to the observation plane. Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit. The WOFRY package in OASYS permits to do this work in a quick way, and it is possible to arrive to a satisfactory result in a few iterations. The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}. The final image profiles are in Fig.~\ref{wofry1D}.

\begin{table}[]
    \centering
    \caption{Beam sizes calculated with simplified wave optics propagation (point source and ideal lenses with aperture for modeling the element dimensions). \todo{implement a sort of multielectron system - implement the slope error} }
    \label{tablewofry1D}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{cccccc} 
    Element:                & ML            & VSS          & KBv               & KBh           & sample        \\
    Position [m]:           & 28.3          & 40.00        & 184.90            & 184.95        & 185.00        \\
    H beam size [$\mu$m]:   & 328           & 1.11         & 4084              & 390           & 0.010         \\
    V beam size [$\mu$m]:   & 328           & 466          & 899               & 450           & 0.010         \\ 
    H scaling factor:       & 1.0           & 0.01         & 220               & 1             & 0.00002       \\
    V scaling factor:       & 1.0           & 0.64         & 5                 & 0.5           & 0.00002       \\
    \end{tabular}%
    }
\end{table}

\begin{figure}
\label{wofry1D}
\centering
\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dh.png}
\includegraphics[width=4.25cm]{GRAPHICS/wofry1Dv.png}
\caption{Intensity distribution at the focal position in horizontal plane (left) and in vertical plane (right).}
\end{figure}

\subsubsection{Full wave optics simulation for coherent case}
% The reduction on the electron beam emittance, by reducing the electron beam horizontal size, leads to an increase on the horizontal lateral coherence of the radiation emitted by an undulator in the symmetry point of a straight section \cite{paper}. Incoherent simulation methods may be insufficient for a full assessment of the beamline design. To accurately account for coherence effects, one has to resort to wave optics simulations (commonly referred to as wave propagation).
% 
The first step in the wave optics simulation is to assume a full coherent source and simulate the propagation of such wavefront through the beamline. The source used on the simulations to follow is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV, which is inherently fully coherent. Once the source is defined, one has to go around on how to simulate the beamline optical elements. Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture; or b) use grazing incidence focusing optics models \cite{Canestrari2014}. On the top of that, one can add the effect of imperfect optics to the simulations, i.e. slope errors.

The computer code chosen for these calculations is the \textsc{srw} \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.

Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A. It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model, confirming the previous ray-tracing results. The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant difference is found, which is in accordance with what is reported in \cite{Canestrari2014}. Notica that the differences are minimal and have to be observed in log plots. It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid at the focal position vicinity: down- or upstream the focal position, differences are not negligible \cite{Canestrari2014}.

It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}). For that, the available available metrology data (Appendix 1) are used. For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}). There is a increase in the background level: the radiation that should be on the main lobe (focus) goes to the side lobes as a result of adding the slope errors. This is accompanied by an obvious reduction on the peak intensity. Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.

>>>>

The next step would be to estimate the number of photons at the sample. For that, we need the spectrum at the source. It is not straightforward to get this information ``by hand'' but one can calculate the flux and power using different codes. Perhaps the most performant and well maintained codes spread in the synchrotron community are SPECTRA \cite{codeSPECTRA} and SRW \cite{codeSRW}. The XOPPY package in the OASYS suite implements three codes: US\cite{codeUS}, URGENT: \cite{codeURGENT} and SRW \cite{codeSRW}. Figure~\ref{fig:FluxU18} shows the spectra calculated over an aperture if 1 $\times$ 1 mm$^2$ placed at $d$=27.1 m. This aperture opening is set to fully accept the central cone at the resonance peak of the harmonic in use ($\max(\Sigma') \times d = 0.73$ mm), with $\Sigma'$ values from in Table~\ref{tab:HandCalculations}. The peak intensity is about  $2.9\cdot10^{14}$ photons/s/0.1{\%}bw for EBS and 1.9 10$^{14}$ for ESRF (see \ref{fig:FluxU18}). One can appreciate how the emittance plays an important role in both peak width and specially in peak amplitude. The number of photons per second after the multilayer monochromator is 2.9 10$^{14}$ for EBS and 1.9 10$^{14}$ for ESRF, considering that the energy bandwidth transmitted by a multilater monochromator is $\Delta E/E \approx 10^{-2}$ and the bandwidth in \ref{fig:FluxU18} is 10$^{-3}$, and supposing at this point ideal reflectivity (100\%) for the multilayers.

\begin{figure}\label{fig:FluxU18}
    \centering
    \includegraphics[width=0.8\textwidth]{GRAPHICS/fluxU18.png}
    \caption{Flux of the U18.3 undulator integrated over a square slit placed at 27.1~m from the ID centre with aperture of 1 and 4 mm.}
\end{figure}

Next we would like to estimate the geometrical transmitivity of the beamline, therefore we have to consider which optical element crop the beam and remove photons. This can be analysed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h=50~\mu$m in horizontal and fully opened in vertical. For ESRF its transmittivity is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their length to a direction perpendicular to the optical axis. Using mirror length of $L_v$=60 mm and $L_h$=26 mm for VFM and HFM, respectively, and incident angle of 15 mrad, the KB is behaving like an aperture of $900 \times 390 \mu m^2$. We get N.A. values of 6.2$\times$21.6 $mu$rad (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmittivity coefficient is the ratio of the accepted N.A. over the beam divergence, resulting 9.5\% (H) and 0.18\% (V)for ESRF and 16.3\% and 0.18\% for EBS. The overall transmission is the product of H and V ratios, therefore 0.2\% for ESRF and 2.95\% for EBS, therefore 3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS. 
\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? }

One can also discuss some simple consequences about the beam coherence. The coherence fraction (CF) \cite{arxivCF} of the undulator beam is the occupation of the lowest coherent mode, as it will be discussed in Section~\ref{level2}. An approximated estimation of it can be obtained using Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas}. The CF in its the horizontal and vertical planes are
\begin{equation}
 CF_{(h,v)} = \frac{\sigma_u \sigma'_u}{\Sigma_{(h,v)} \Sigma'_{(h,v)}},
\end{equation}
and the total coherent fraction is the product of both (CF=CF$_h \times $CF$_v$). An estimation of the CF calculated in such a mode give CF=0.12\% for the ESRF and 2.73\% for EBS at 17 keV photon energy. The CF roughly indicates the ratio of the ``coherent photons'' over the total number of photons. It is a good indicator of the quality of a coherent beam (a fully coherent beam has CF=1 and an incoherent beam CF=0). The beamline optics acts as a ``coherence filter'' by removing the ``bad'' or ``incoherent'' photons and in consequence increasing the CF. A complete cleaning is impossible because of the spatial overlapping of the different coherence modes. However, apertures centered at the optical axis would make the work. If we compare the values transmittivity of the beamline due to geometrical clipping  (0.2\% for ESRF and 2.95\% for EBS) with the the CF values, we remark that they are very close. This implies that the beam at the focal position would be highly coherent because the fraction of the photons removed from the beam matches the CF value that quantify the fraction of ``good'' or ``coherent'' photons.

To assess the focused beam size, one can idealize the beamline in two parts. First, the ML monochromator and VSS that will prepare a coherent plane wave as the source is very far from the sample position. Second, the KB has a double role, acting as a focusing element of focal distance $f$=5 cm, but also crops the beam as as it would be done by a square aperture of $D \times D=$ 400$\times$400 $\mu m^2$ (the length of the mirror projected in a plane perpendicular to the beam propagation) \inred{(we obtained 900 in H, that will be reduced by 0.5 as propagated from 10 cm to 5 cm upstrem from the sample, and 390 in V approximated by 400)}. This aperture has two effects: i) as discussed before, it removes a lot of off-axis photons increasing the CF to a value close than one, and ii) make a broadening of the focal spot because of coherent diffraction. We know from optics that the diffraction by a square aperture of a collapsing spherical wavefront has an intensity distribution proportional to $\sinc^2(k D x/2 f) \sinc^2(k D y / 2 f)$, where $k=2\pi/\lambda$, and $x$ ($y$) is the horizontal (vertical) coordinate . Considering that the FWHM of $\sinc(x)$ is approximately 2.78, one obtains a FWHM of 7.75$\times$7.75 nm$^2$ for the focal spot. This spot size is of the order or larger than the one obtained by geometrical considerations, therefore contributing significantly to the size of the focused beam. One can conclude that for the conditions in use this is a diffraction limited beamline.

In summary, we show in this section how simple analytical calculations that can be done in its major part ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5 nm$^2$ for ESRF, 10$\times$5 nm$^2$ for the EBS) the flux (3.8 $10^{11}$ photons/s for ESRF and 8.5 $10^{12}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction limited focal spot of  7.75$\times$7.75 nm$^2$. We ignored at this point the effect of real optics containing surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. It also helps in having a critical eye or sanity check on the numerical calcuations that will provide more precised values, but must remain the this order of magnitude.

@@

In this section we perform ray tracing calculations to study the effect of real optics (including aberrations and slope errors) to the beam size and beamline transmittivity. The first subsection deals with the effect of aberrations and transmittivity due to geometrical considerations. The second subsection shows how the effect of slopes errors the KB mirrors degrade the focal spot.

Ray tracing is a simulation method based on tracing some light rays along the optical system from the source to the image and retrieve the statistics of rays at the image to measure beam characteristics. For simple systems one can perform ray tracing by hand, selecting some principal rays that will define the location and envelope of the image. Using computers one can trace thousands of rays and calculate the position and divergence distribution of the rays using statistics. Rays are usually generated by Monte Carlo sampling the source characteristics. Ray tracing is based on geometrical optics: a ray is a solution of the Helmholtz equation and travels in vacuum along straight line. Rays are intercepted by the optical elements, which change their trajectory. For reflectors, this change of direction is given by the specular reflection laws. For refractors, the change of direction os defined by the Snell law. Because Snell law uses the refraction index which is wavelength-dependent, it is then interetisting to add additional parameters to each ray, such as the wavelength. This will permit to calculate the refraction index that has to be used by each ray. The information of each ray can also be extended to include some intensity, or better electric field amplitude. This allows to extend the pure geometrical tracing to include physical models that take into account the optical element reflectivity or transmittivity. This combination of using a geometrical model plus a physical model allows to simulate every element used in a synchrotron beamline, such as mirrors and combination of them (e.g., KBs); lenses, compound refractive lenses and transfocators; gratings of any type and shape, crystal systems, etc. Ray tracing is a simple and extremely powerful technique for calculating the main characteristics of the photon beam (size, divergence, photon distribution) at every point of the beamline. Although very powerful commercial ray tracing packages are available in the market, they are not suitable for synchrotron simulations because they do not cover the particularities of X-ray devices and sources. For that purposes, several packages created in the synchrotron community and are available, like SHADOW \cite{codeSHADOW}, RAY \cite{codeRAY}, McXtrace \cite{codeMCXTRACE} or XRT \cite{codeXRT}. SHADOW is probably the one that is more used and cited. We use SHADOW and its interface ShadowOUI \cite{codeSHADOWOUI}, available as a module of the OASYS suite.

@@

We performed ray tracing using the elliptical reflectors with finite dimensions as focusing elements. The reflector shape is chosen to be elliptical as this is the theoretical shape producing a point-to-point focusing. For a first calculation  the elliptical elliptical reflectors are considered perfect, and have no slope errors. 
The results in Fig.~\ref{fig:ray tracing}a show two facts. One is that the migration from ESRF lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF. Secondly, one can observe a much larger luminosity in the EBS case. The transimittivity values are 0.165\% for ESRF and 2.28\% for EBS. These values confirm what we calculated by hand (0.2\% for ESRF and and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{coherentfraction} (0.12 for ESRF and 2.73\% for EBS), meaning that a very high coherence of the beam is expected, therefore some degradation of the beam by clipping the beam by apertures and element acceptance is expected.

@@

The surface finish of the optical elements is another limiting effect for determining the size and intensity of the beam at the image position. It is possible to infere the effect of mirror imperfections and in particular slope errors by only knowing some statistical parameters such as the slope error Root Mean Squate (RMS) and then building synthetic surface profiles to be used in the simulations. This is the typical mode to operate when designing a new beamline where the mirrors do not exist yet. However, in our case the mirrors already exist and have been extensively characterize at the ESRF metrology laboratory. It is therefore highly recommended to use metrology data from real mirrors, when possible. When not available, one can use data from existing mirrors from a database (e.g., \cite{dabam}) and extrapolate some of the effects to the hypothetical new mirrors. The metrology data of the mirrors for the beamline under study are presented in \todo{Appendix 1}. These experimental data are used in all the simulations in this paper. Simulations using the hybrid model including slopes errors are shown in Fig.~\ref{fig:ray tracing}c. It can be observed a small broadening of the peak due to slope errors, but, more importantly, a new satellite structure close to the vertical intensity peak is produced by the slope errors.

\todo{
\begin{itemize}
 \item play with slope errors and use synthetic profiles to show the importance of the low frequencies in the intensity distribution?
 \item explain why KB errors are treated with hybrid and not with shadow. In Discussion section? 
\end{itemize}
}

@@

In this section the coherence properties of the beamline are analysed using, again, different methodologies classified by complexity and computer resources. We first analyze the beamline supposing a fully coherent beam. Subection~\ref{wofry} shows a very simpified 1D model. Subsection~\ref{srw_se} used a full 2D model with accurate simulation of the undulator emission but without considering storage ring emittance. These simplifications permit to address diffraction effects by apertures and slope errors. Then we introduce methods to treat the partial coherence originated by the storage ring emittance. Two variants are shown, a Monte Carlo approach (subsection~\ref{srw_me}) that permits us to calculate in great detail the intensity dostribution of the focused beam, and a decomposition of the radiation in coherent modes (subsection~\ref{comsyl}) that allows to evaluate quantitatively the quality of the coherent beam.

\subsubsection{Simplified wave optics simulation (coherent case in 1D with point source and ideal elements)}
\label{wofry}

The very long distance source-sample combined with a small source (in particular for EBS) makes this beamline very adapted for studied requiring coherence. Indeed, it has been shown that the clipping of the beam, mainly due to the acceptance of the KB mirrors produces a significant broadening of the focal spot.  As discussed in Section~\ref{level0}, very simple calculation of the diffraction limit can be sone using a convergent spherical beam clipped by a slit of 400$\times$400$ \mu$m$^2$ placed at 5cm upstream the focus (at the second KB mirror position). The slit vertical size corresponds roughly to the horizontal acceptance of the seconf KB mirror, and it is approximately the same in vertical. A simple simulation of diffraction by a square slit using OASYS is shown in Fig.~\ref{wofry400um}. It reveals a spot of less than 10 nm, thus confirming the analytical predictions, and qualifying this beamline to be a diffraction limited beamline.

\begin{figure}
\label{wofry400um}
\centering
\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umslit.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry400umprofile.png}
\caption{Diffraction spot producedof a spherical wavefront (R=-50mm) by a 400x}
\end{figure}

We perform here a simplified wavefront simulation in 1D, thus separating the simulations of the horizontal and vertical planes. The source can be simplified by using a point source (spherical wavefront) and applying a Gaussian intensity profile with $\sigma=\sigma'_u d$ with $\sigma'_u$ given by Eq.~\ref{eq:photon small sigmas} and $d$ the distance from the point source to the observation plane. Each optical element can be simplified by separating its action in two parts: the focusing behaviour implemented as an ideal lens, and the finite size because of the boundaries, implemented as a slit. The WOFRY package in OASYS permits to do this work, and it is possible to arrive to a satisfactory result in a few iterations. The beam size at different positions of the beamline are shown in Table~\ref{tablewofry1D}. The final image profiles are in Fig.~\ref{wofry1D}.

\begin{table}[]
    \centering
    \caption{Beam sizes calculated with simplified wave optics propagation (point source and ideal lenses with aperture for modeling the element dimensions). 
    }
    \label{tablewofry1D}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{cccccc} 
    % using WAVEOPTICS WOFRY workspace
    Element:                & ML            & VSS          & KBv               & KBh           & sample        \\
    Position [m]:           & 28.3          & 40.00        & 184.90            & 184.95        & 185.00        \\
    H beam size [$\mu$m]:   & 329           & 1.11         & 4084              & 390           & 0.008         \\
    V beam size [$\mu$m]:   & 329           & 466          & 899               & 450           & 0.007         \\ 
    H scaling factor:       & 1.0           & 0.01         & 440               & 1             & 0.00002       \\
    V scaling factor:       & 1.0           & 0.64         & 5                 & 0.5           & 0.00002       \\
    \end{tabular}%
    }
\end{table}

\begin{figure}
\label{wofry1D}
\centering
\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dh.png}
\includegraphics[width=0.45\textwidth]{GRAPHICS/wofry1Dv.png}
\caption{Intensity distribution at the focal position in horizontal plane (left, FWHM=8.1 nm) and in vertical plane (right, FWHM=7.1 nm).
\todo{implement a sort of multielectron system - implement the slope error. In Discussion? } 
}
\end{figure}

\subsubsection{Full wave optics simulation for coherent case}
\label{srw_se}
% The reduction on the electron beam emittance, by reducing the electron beam horizontal size, leads to an increase on the horizontal lateral coherence of the radiation emitted by an undulator in the symmetry point of a straight section \cite{paper}. Incoherent simulation methods may be insufficient for a full assessment of the beamline design. To accurately account for coherence effects, one has to resort to wave optics simulations (commonly referred to as wave propagation).
% 
A next step in the wave optics simulation is, maintaining the full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that match well the characteristics of the real elements. The source simulated is the on-axis-single-electron emission from the U18.3 undulator tuned to its first harmonic at $E=17225$ eV. The emission by a single-electron is fully coherent. Once the source is defined, one has to go around on how to simulate the beamline optical elements. Increasing the complexity of the simulations one can: a) use ideal thin lenses preceded by slits to emulate finite aperture (as done in the 1D model presented in Section~\ref{wofry}); or b) use grazing incidence focusing optics models \cite{Canestrari2014}. On the top of that, one can add the effect of imperfect optics to the simulations, i.e. slope errors.

The computer code chosen for these calculations is Synchrotron Radiation Workshop (\textsc{srw}) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.

Figure \ref{fig:SingleElectron} shows the horizontal integrated intensity at the focal position for the ID16A. It can been seen that there is little difference between using thin ideal lenses or the elliptical mirror model. The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant difference is found, which is in accordance with what is reported in \cite{Canestrari2014}. Notice that the differences are minimal and to remark them we have chosen log plots to represent the intensity distributions at the focal position. It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid at the focal position vicinity: down- or upstream the focal position, differences are not negligible \cite{Canestrari2014}.

It is also possible to add slope errors to the simulations to study the degradation on the focal spot quality (Fig.~\ref{fig:SingleElectron}). For that, the available available metrology data \inred{(Appendix 1)} are used. For both cases (i.e., simulating the beamline using thin optical elements or elliptical mirrors), the beam sizes calculated as the FWHM at the focal position are little affected by the presence of the slope errors (see Table \ref{tab:SRW_results}). There is a significant reduction and shuffling of the interference fringes observed with no slope error. They are almost cancelled, with is an important raise of the background level: the radiation found on the main lobe (focus) shifts to the side lobes when the slope errors are considered. This is accompanied by an obvious reduction on the peak intensity. Table \ref{tab:SRW_results} compiles the beam sizes at focal position for mentioned simulations.

\begin{figure}
    \centering
        \includegraphics[width=4.25cm]{GRAPHICS/se_idealTE_idealOE_h.pdf}
        \includegraphics[width=4.25cm]{GRAPHICS/ebs_slopeTE_slopeOE_h.pdf}
    \label{fig:SingleElectron}
    \caption{Wavefront propagation from a single electron emission through the ID16A beamline. Left: the perfect focusing with ideal lenses (black) is compared with the focusing using elliptical mirrors (red). Right: the same as in the previous plot, but adding the effect of the mirrors slope errors in bothe cases.}
    %Simulations took less than one minute in a desktop computer Intel Core i7 with 8Gb of RAM memory.}
\end{figure}
